---
title: "Final Project"
author: "Connor Lawson - Darrious Barger"
date: "5/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(resampledata)
library(knitr)
library(dplyr)
library(skimr)
library(utils)
library(pls)
library(ISLR)
library(readxl)
library(MASS)
library(boot)
library(caret)
library(glmnet)
library(leaps)
```



```{r, include = FALSE}
housing <- read_excel("Housing.xlsx")
```


# Model from the previous project

```{r}
model1 = lm(price ~ garagesize + size + bedrooms + I(bedrooms^2), data = housing)
summary(model1)
```

# Using regsubsets on all the data


```{r}
housing$elem <- as.numeric(c("edison" = 0, "adams" = 1, "parker" = 2, "edge" = 3, "harris" = 4, "crest" = 5)[housing$elem])
housing$status <- as.numeric(c("sld" = 0, "pen" = 1, "act" = 2)[housing$status])
```

```{r}
fow_sel <- regsubsets(price ~ ., data = housing, nvmax = 11, method = "forward")
regSum <- summary(fow_sel)
regSum
```

```{r}
regSum$rsq
regSum$adjr2
regSum$cp
regSum$bic
```


```{r}
which.max(regSum$rsq)
which.max(regSum$adjr2)
which.min(regSum$cp)
which.min(regSum$bic)
```


```{r}
coef(fow_sel,7)
```


```{r}
model2 <- lm(price ~ id + size + bath + bedrooms + agestandardized + status + elem, data = housing)
summary(model2)
```

# Creating training and test data

```{r}
set.seed(123)
half <- sample(nrow(housing), nrow(housing) * 0.5, replace = FALSE)
train <- housing[half, ]
test <- housing[-half, ]
```

# Using regsubsets on the training data

```{r}
test.mat <- model.matrix(price ~ ., data = test)
```

```{r}
regfit.best <- regsubsets(price ~ ., data = train, nvmax = 11, method = "forward")
```



```{r}
val.errors <- rep(NA, 11)
```

```{r}
for (i in 1:9) 
{
  coefi=coef(regfit.best,id=i)
  pred=test.mat[,names(coefi)]%*%coefi
  val.errors[i] <- mean((test$price-pred)^2)
}
val.errors
```

```{r}
which.min(val.errors)
```

```{r}
regfit.best2 <- regsubsets(price ~ ., data = train, nvmax = 4, method = "forward")
summary(regfit.best2)
```

A model using 5 predictors will be the best.

```{r}
model3 <- lm(price ~ size + bedrooms + elem + lot + garagesize, data = housing )
summary(model3)
```


# Using Ridge Regression


```{r}
set.seed(123)

x=model.matrix(price ~ ., data = train)[,-1]
y=train$price
```


Below, we are setting up the lamba values that we will test. We are also using the `glmnet()` function to fit the ridge regression.

```{r}
set.seed(123)

grid=10^seq(10,-2,length=100)
ridge.mod=glmnet(x,y,alpha=0,lambda=grid)
```


This finds the value of lambda at the 50th iteration of the sequence

```{r}
ridge.mod$lambda[50]
```


```{r}
coef(ridge.mod)[,50]
```


This will calculate the sum of squared betas in the ridge regression model.

```{r}
sqrt(sum(coef(ridge.mod)[-1,50]^2))
```


The predict function will give us the ridge regression coefficients for lamba = 50.

```{r}
predict(ridge.mod,s=50,type="coefficients")
```

We use cross Validation to find the best lamda.

```{r}
set.seed(1)
train=sample(1:nrow(x), nrow(x)/2)
test=(-train)
y.test=y[test]
ridge.mod=glmnet(x[train,],y[train],alpha=0,lambda=grid, thresh=1e-12)

ridge.pred=predict(ridge.mod,s=4,newx=x[test,])
```


# Partial Least Squares

```{r}
set.seed(123)

pls.fit=plsr(price ~ .,data = housing, subset = train, scale=TRUE, validation="CV")

summary(pls.fit)
```

```{r}
validationplot(pls.fit,val.type="MSEP")
```

Looking at the graph, we will choose a 1 component model since the graph is growing larger and we would like to minimize the error.


```{r}
pls.pred = predict(pls.fit, newx = x[train,], ncomp = 1)
pls.fit = plsr(price ~ ., data = housing,scale=TRUE,ncomp = 1)
```

# Calculate the MSE

MSE For model 1

```{r}
mean((housing$price - predict(model1, housing))^2)
```

MSE for model 2

```{r}
mean((housing$price - predict(model2, housing))^2)
```

MSE for model 3


```{r}
mean((housing$price - predict(model3, housing))^2)
```


MSE for model 4

```{r}
mean((ridge.pred-y.test)^2)
```


MSE for model 5

```{r}
mean((pls.pred-y.test)^2)
```

